Deeply Integrating C11 Code Support into Isabelle/PIDE

### 2019-10-17 C11-FrontEnd/document/generated/paper.tex
synchronize with fortesse
synchronize with fortesse revision 5374

%
\begin{isabellebody}%
\setisabellecontext{paper}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimML
%
\endisadelimML
%
\isatagML
%
\endisatagML
{\isafoldML}%
%
\isadelimML
%
\endisadelimML
%
\begin{isamarkuptitle*}%
[label = {tit},type = {scholarly_paper.title}, args={label = {tit},type = {scholarly_paper.title}, scholarly_paper.title.short_title = {}}]Isabelle/C%
\end{isamarkuptitle*}\isamarkuptrue%
%
\begin{isamarkuptext*}%
[label = {abs},type = {scholarly_paper.abstract}, args={label = {abs},type = {scholarly_paper.abstract}, scholarly_paper.abstract.keywordlist = {{User Interface, Integrated Development, Program Verification, Shallow Embedding}}}]We present a framework for C code in C11 syntax deeply integrated into the Isabelle/PIDE
  development environment. Our framework provides an abstract interface for verification back-ends
  to be plugged-in independently. Thus, various techniques such as deductive program verification or
  white-box testing can be applied to the same source, which is part of an integrated PIDE document
  model. Semantic back-ends are free to choose the supported C fragment and its semantics. In
  particular, they can differ on the chosen memory model or the specification mechanism
  for framing conditions.

@@

Orca \cite{bockenek:hal-02069705}, or Clean (discussed in this paper). A particular advantage of

@@

Note that the C11 grammar has enough scoping structure for the full inference of the C environment
\isatt{env} be at bottom-up time. In terms of efficiency, we use specific
\emph{static} rule wrappers having the potential of overloading default grammar rules (see \csname isaDof.ref\endcsname[type={Isa_COL.figure}]{architecture}), to assign a wrapper to be always executed as soon as a Shift-Reduce
rule node of interest is encountered. The advantage of this construction is that the wrappers are
statically compiled, which results in a very efficient reporting of C type information.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkupsection*}%
[label = {backends},type = {scholarly_paper.technical}, args={label = {backends},type = {scholarly_paper.technical}, scholarly_paper.text_section.main_author = {@{docitem ''bu''}}, Isa_COL.text_element.level = {}, Isa_COL.text_element.referentiable = {False}, Isa_COL.text_element.variants = {{STR ''outline'', STR ''document''}}, scholarly_paper.text_section.fixme_list = {}, Isa_COL.text_element.level = {}, scholarly_paper.technical.definition_list = {}}]Semantic Back-Ends%
\end{isamarkupsection*}\isamarkuptrue%
%
\begin{isamarkuptext}%
In this section, we briefly present two integrations of verification back-ends for C. We
chose Clean~\footnote{stemming from recent HOL-TestGen distributions
    \url{https://www.brucker.ch/projects/hol-testgen/index.en.html}} used
for program-based test generation \cite{DBLP:conf/tap/Keller18}, and AutoCorres \cite{DBLP:conf/pldi/GreenawayLAK14}, arguably the most developed deductive verification environment
for machine-oriented C available at present.

Note that we were focusing on keeping modifications of integrated components minimal, particularly
for the case of AutoCorres. Certain functionalities like position propagation of HOL terms in
annotations, or ``automatic'' incremental declarations~\footnote{\url{https://github.com/seL4/l4v/blob/master/tools/autocorres/tests/examples/Incremental.thy}}
may require internal revisions on the back-end side. This is out of the scope of this paper.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\begin{figure}
  \centering
  \begin{minipage}{0.53\linewidth}
  \includegraphics[width=0.96\textwidth]{figures/A-C-Source7}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
\begin{isar}
prime$_{\text{C}}$_core_def: "prime$_{\text{C}}$_core n \<equiv>
  if$_{\text{Clean}}$ \<Open> (n < 2) \<Close> then return 0 else skip;-
  \<Open> i := 2 \<Close>;-
  while$_{\text{Clean}}$ \<Open> i < SQRT_UINT_MAX \<and> i * i \<le> n\<Close>
    (if$_{\text{Clean}}$ \<Open>n mod i = 0\<Close>
      then return 0 else skip;
     \<Open>k:=k+1\<Close>; assert \<Open> k\<le>UINT_MAX \<Close>
     \<Open>i:=i+1\<Close>; assert \<Open> i\<le>UINT_MAX \<Close>) ;-
  return 1"

@@

%% \end{wrapfigure}%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{paper}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimML
%
\endisadelimML
%
\isatagML
%
\endisatagML
{\isafoldML}%
%
\isadelimML
%
\endisadelimML
%
\begin{isamarkuptext*}%
[label = {abs},type = {scholarly_paper.abstract}, args={label = {abs},type = {scholarly_paper.abstract}, scholarly_paper.abstract.keywordlist = {{User Interface, Integrated Development, Program Verification, Shallow Embedding}}}]We present a framework for C code in C11 syntax deeply integrated into the Isabelle/PIDE
  development environment. Our framework provides an abstract interface for verification back-ends
  to be plugged-in independently. Thus, various techniques such as deductive program verification or
  white-box testing can be applied to the same source, which is part of an integrated PIDE document
  model. Semantic back-ends are free to choose the supported C fragment and its semantics. In
  particular, they can differ on the chosen memory model or the specification mechanism
  for framing conditions.

@@

Orca \cite{bockenek:hal-02069705}, or Clean \cite{journals/afp/TuongW19}. A particular advantage of

@@

Note that the C11 grammar has enough scoping structure for the full inference of the C environment
\isatt{env} be at bottom-up time. In terms of efficiency, we use specific
\emph{static} rule wrappers having the potential of overloading default grammar rules (see \csname isaDof.ref\endcsname[type={Isa_COL.figure}]{architecture}), to assign a wrapper to be always executed as soon as a Shift-Reduce
rule node of interest is encountered. The advantage of this construction is that the wrappers are
statically compiled, which results in a very efficient reporting of C type information.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkupsection*}%
[label = {backends},type = {scholarly_paper.technical}, args={label = {backends},type = {scholarly_paper.technical}, scholarly_paper.text_section.main_author = {@{docitem ''bu''}}, Isa_COL.text_element.level = {}, Isa_COL.text_element.referentiable = {False}, Isa_COL.text_element.variants = {{STR ''outline'', STR ''document''}}, scholarly_paper.text_section.fixme_list = {}, Isa_COL.text_element.level = {}, scholarly_paper.technical.definition_list = {}}]Semantic Back-Ends%
\end{isamarkupsection*}\isamarkuptrue%
%
\begin{isamarkuptext}%
In this section, we briefly present two integrations of verification back-ends for C. We
chose Clean \cite{journals/afp/TuongW19}~\footnote{stemming from recent HOL-TestGen distributions
    \url{https://www.brucker.ch/projects/hol-testgen/index.en.html}} used
for program-based test generation \cite{DBLP:conf/tap/Keller18}, and AutoCorres \cite{DBLP:conf/pldi/GreenawayLAK14}, arguably the most developed deductive verification environment
for machine-oriented C available at present.

Note that we were focusing on keeping modifications of integrated components minimal, particularly
for the case of AutoCorres. Certain functionalities like position propagation of HOL terms in
annotations, or ``automatic'' incremental declarations~\footnote{\url{https://github.com/seL4/l4v/blob/master/tools/autocorres/tests/examples/Incremental.thy}}
may require internal revisions on the back-end side. This is out of the scope of this paper.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkupsubsection*}%
[label = {clean},type = {scholarly_paper.technical}, args={label = {clean},type = {scholarly_paper.technical}, scholarly_paper.text_section.main_author = {@{docitem ''bu''}}, Isa_COL.text_element.level = {}, Isa_COL.text_element.referentiable = {False}, Isa_COL.text_element.variants = {{STR ''outline'', STR ''document''}}, scholarly_paper.text_section.fixme_list = {}, Isa_COL.text_element.level = {}, scholarly_paper.technical.definition_list = {}}]A Simple Typed Memory Model: Clean%
\end{isamarkupsubsection*}\isamarkuptrue%
%
\begin{isamarkuptext}%
Clean (pronounced as: ``C lean'' or ``Céline'' [selin]) is based on a simple,
shallow-style execution model for an imperative target language.
Importing Clean into a theory with its activated back-end proceeds as in \autoref{fig:clean}.
For a deep presentation of Clean, the reader is referred to the published AFP document \cite{journals/afp/TuongW19}.

\begin{figure}
  \centering
  \begin{minipage}{0.53\linewidth}
  \includegraphics[width=0.96\textwidth]{figures/A-C-Source7}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
\begin{isar}
prime$_{\text{C}}$_core_def: "prime$_{\text{C}}$_core n \<equiv>
  if$_{\text{Clean}}$ \<Open> (n < 2) \<Close> then return 0 else skip;-
  \<Open> i := 2 \<Close>;-
  while$_{\text{Clean}}$ \<Open> i < SQRT_UINT_MAX \<and> i * i \<le> n\<Close>
    (if$_{\text{Clean}}$ \<Open>n mod i = 0\<Close>
      then return 0 else skip;
     \<Open>k:=k+1\<Close>; assert \<Open> k\<le>UINT_MAX \<Close>
     \<Open>i:=i+1\<Close>; assert \<Open> i\<le>UINT_MAX \<Close>) ;-
  return 1"

@@

Finally, we will have a glance at the code for the registration of the annotation commands presented
in \csname isaDof.ref\endcsname[type={Isa_COL.figure}]{C-sample10}, and used in the example of \autoref{fig:clean}. Thanks to
Isabelle/C's function \isatt{C{\char`\_}Annotation.command{\char`\'}}, the registration of
user-defined annotations is very similar to the registration of ordinary commands in the Isabelle
platform.%
